---
title: "Personal activity monitoring device"
author: "L. Besnainou"
date: "19 février 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(DT)
knitr::opts_chunk$set(echo = TRUE)
```



```{r readingFile}
# Reading the data
activity <- read.csv("activity.csv",na.strings = "NA")
```

## Histogram of the total number of steps taken each day

```{r First_Plot}
#Transform the date column in a valide date format
activity$date <- ymd(activity$date)
#Make the histogram plot in ggplot2
activity %>% group_by(date) %>% summarize(numb=sum(steps,na.rm = TRUE)) %>%
    ggplot(aes(x=date,y=numb))+
    geom_col(aes(colour = "tomato"),show.legend=F)+
    ggtitle("Histogram of the total number of steps taken each day")

```

## Calculate the Mean and median number of steps taken each day
as a lot of measurements of steps are 0 the median value is always 0
```{r Mean and Median}
table <- activity %>% filter(!is.na(steps)) %>% group_by(date) %>% 
    summarise(mean=mean(steps,na.rm = TRUE),median=median(steps,na.rm = TRUE))
                                                                             
datatable(table)
```


## Time series plot of the average number of steps taken
```{r, timeserieplot}
 table %>% ggplot(aes(date, mean)) + geom_line(colour = "red") + ggtitle ("Average number of steps taken")

```

##The 5-minute interval that, on average, contains the maximum number of steps

```{r}
IntervalAvg <- activity %>% group_by(interval) %>% summarise(mean=mean(steps,na.rm = T))
MaxInterval <- IntervalAvg[max(IntervalAvg$mean)==IntervalAvg$mean,]
```
The interval is <span style="color:red">`r MaxInterval[1,1]`</span> with an average number of steps <span style="color:red">`r MaxInterval[1,2]`</span>

##Code to describe and show a strategy for imputing missing data

The missing values for the steps will be replace by the average value  for the same interval (calculated in IntervalAvg R variable)
```{r}
#A left join will be made on dataframe Activity
activityNew <- left_join(activity,IntervalAvg,by = c("interval"))
activityNew$steps <-  ifelse(is.na(activityNew$steps), activityNew$mean, activityNew$steps)

```


## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
activityNew %>% group_by(date) %>% summarize(numb=sum(steps,na.rm = TRUE)) %>%
    ggplot(aes(x=date,y=numb))+
    geom_col(aes(colour = "tomato",fill="tomato"),show.legend=F)+
    ggtitle("Histogram of the total number of steps taken each day after missing values are imputed")

```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r facet,fig.width = 14, fig.height = 9, out.width = "800", out.height = "600"}
#set weekday and week-end
activity$weekday <- weekdays(activity$date)
activity$wday <- wday(activity$date) %in% c(1,7)
activity$wday <-factor (activity$wday)
levels(activity$wday) <- c("Semaine", "week-end")
p <- activity %>% group_by(wday,weekday) %>% summarise(mean=mean(steps,na.rm = T)) %>%
ggplot(aes(weekday,mean,fill=wday)) +
  geom_col( aes(colour = "tomato"),show.legend=F)+scale_x_discrete(limits=c("samedi", "dimanche",  "lundi", "mardi", "mercredi","jeudi", "vendredi" ))
p +  facet_wrap(~ wday)
```


